\LARGE \bf Enhancing Security in Multi-Robot Systems through Co-Observation Planning on Unsecured Trajectories

### 2023-08-02 paper.tex
Updates from Overleaf

\section{Preliminaries}
In this section, we review various mathematical concepts that will provide the foundations and context for our novel constraint.
\subsection{Ellipsoidal reachability constraints}\label{Chapter:ProblemFormulation}
  The reachability region is defined as the set of locations $x(t)$ that a robot can reach between two given fixed positions:

@@

{\small
\bibliographystyle{unsrt}
\bibliography{ADMM_planning,ACC}
}

>>>>

\section{Preliminaries}
In this section, we review various mathematical concepts that will provide the foundations and context for our novel constraint.

In this section, we review various mathematical concepts that will provide the foundations and context for our novel constraints.

\subsection{Differentials}
We define the differential of a map $f(x):\real{m}\to\real{n}$ at a point $x_0$ as the unique matrix $\partial_x f \in\real{n\times m}$ such that
\begin{equation}\label{equ:dt_to_dx}
  \left.\dert f\bigl(x(t)\bigr)\right|_{t=0}=\partial_x f\bigl(x(0)\bigr) \dot{x}(0)
\end{equation}
where $t\mapsto x(t)\in\real{n}$ is a smooth parametric curve such that $x(0)=x$ with any arbitrary tangent $\dot{x}(0)$. For later part of this paper, we will use $\dot f$ for $\dert f$ and $\partial_x f$ for $\frac{\partial f}{\partial x}$. The differentials $\partial_x f$ is derived through (\ref{equ:dt_to_dx}) having $\dot f$ divided by $\dot x$.

With a slight abuse of notation, we use the same notation $\partial_xf$ for the differential of a matrix-valued function with scalar arguments $f:\real{R}\to\real{m\times n}$  Note that in this case \eqref{equ:dt_to_dx} is still formally correct, although semantically different.

\subsection{Alternating Directions Method of Multipliers (ADMM)}\label{chapter:ADMM review}
In this section, we briefly review the path planning algorithm based on ADMM constraints discussed in \cite{yangmulti}.
This algorithm can deal with a variety of types of non-convex and non-smooth optimization constraints that are commonly encountered in trajectory optimization tasks once they are formulated in a certain format.
The goal of the section is to lay the groundwork for proposing our novel constraint in the following sections.

Let $x_{ij}\in\real{m}$ denote the position of agent $i\in\{1,\ldots,n\}$ at the discrete-time index $j\in[0,\ldots,T]$, with $m$ being the dimension of the workspace, $n$ the number of agents, and $T$ the planning time horizon. Let $\vx=\stack(\{x_{ij}\})$ denote the aggregated vector of all the agents' trajectories over the entire time interval. The goal of the algorithm is to optimize an objective function $\varPhi(\vx)$ on a set $\Omega$ defined by nonlinear constraints. To solve the problem using ADMM, we introduce a new set of variables $\vz=D(\vx)$ to translate the constraints $\Omega$ on $x$ to (typically lower-dimensional) sets of constraints on $\vz$, such that the projection operator on $\zeta$ can be easily implemented (e.g., a constrained distance between two agents is translated into their difference constrained to an origin-centered sphere).
The new set of constraints $\zeta$ on $z$ is incorporated using an indicator function $\varTheta_\zeta$. The main problem is written as:
\begin{equation}\label{eq:ADMMSetConstraint_general}
		\begin{array}{r@{\quad}c}
    	\min&\varPhi(\vx)+\varTheta_\zeta(\vz) \\
        s.t.& D(\vx)-\vz=0,
      \end{array}
  \end{equation}
  where $D(\vx)= [D_1(\vx)^T,\dots,D_l(\vx)^T]^T$ is a vertical concatenation of different functions for different constraints.

  And having the update steps shown as:
  \begin{subequations}\label{eq:ADMMupdate}
    \begin{align}
      \label{eq:admm-mod-update-q}\vx^{k+1}&:=\argmin_\vx(\varPhi(\vx^k) +\frac{\rho}{2}\norm{D(\vx)-\vz^k+\vu^k}_2^2)\\
      \vz^{k+1}&:=\Pi_\zeta(D(\vx^{k+1})+\vu^k) \label{eq:ADMM_constraint_update}\\
      \vu^{k+1}&:=\vu^k+D(\vx^{k+1})-\vz^{k+1},
    \end{align}
  \end{subequations}
  Previous work in \cite{yangmulti} considered the following types of constraints:
  \begin{lenumerate}{C}
  \item \emph{Velocity constraints}: The movement of each agent is constrained by a maximum distance in any direction over a single discrete time step. This constraint could be adapted also to enforce more refined dynamical models.
  \item \emph{Convex obstacles}: We use convex polygons to model regions defining the boundary of the workspace, as well as solid obstacles that cannot be entered by agents (note that the latter represent a non-convex constraint set from the point of view of the optimization problem). Obstacles with non-convex shapes can be modeled using unions of the convex sets.
  \item \emph{Waypoints with flexible deadlines}: We give locations that the agents need to visit at any point in a given time window (which can potentially contain a single time). These constraints can also be interpreted as static, trusted cameras that can complement the introspection constraints.
  \item\label{it:introspection} \emph{Introspection constraints}: In a given time window, agents are asked to remain at a certain distance from each other to detect each other's presence.
  \end{lenumerate}
  Constraint~\ref{it:introspection} is the main mechanism that was provided in~\cite{yangmulti} for forcing agents to stay away from the forbidden region. However, this mechanism alone cannot provide strong guarantees in this regard. This motivates the introduction of our novel \emph{ellipsoidal constraints} in Sections~\ref{sec:ellipsoid-point},~\ref{sec:ellipsoide-plane}. This novel constraint is defined with the assumption that the velocity constraints still hold under attack.

  \subsection{Householder rotations}\label{sec:householder}

  To formulate the operator $\Pi_\zeta$ for our novel constraint, we will define a differentiable transformation of the constraint set to a canonical form. This transformation will include a rotation that we derive from a modified version of Householder transformations \cite{householder1958unitary}. With respect to the standard definition, our modification ensures that the final operator is a proper rotation (i.e., not a reflection). We call our version of the operator a \emph{Householder rotation}. In this section we derive Householder rotations and their differentials for the 3-D case; the 2-D case can be easily obtained by embedding it in the $z=0$ plane.

  \begin{definition} Let $\nu_\cF$ and $\nu_\cE$ be two unitary vectors ($\norm{\nu_\cF}=\norm{\nu_\cE}=1$). Define the normalized vector $u$ as
    \begin{equation}
      u'=\nu_\cF+\nu_\cE,\quad u =\frac{u'}{\norm{u'}}.
    \end{equation}
    The \emph{Householder rotation} $H(\nu_\cF,\nu_\cE)$ is defined as
    \begin{equation}\label{eq:H definition}
      H(\nu_\cF,\nu_\cE) = 2 u u\transpose-I.
    \end{equation}
  \end{definition}
  The main property of interest for our application is the fact that $H$ is a rotation mapping $\hat{\nu_\cF}$ to $\hat{\nu_\cE}$, as shown by the following.
  \begin{proposition}
    The matrix $H$ has the following properties:
    \begin{enumerate}
    \item It is a rotation, i.e.
      \begin{enumerate}
      \item\label{it:orthonormality} $H\transpose H=I$;
      \item\label{it:determinant} $\det(H)=1$.
      \end{enumerate}
    \item\label{it:transformation} $\nu_\cE=H \nu_\cF$.
    \end{enumerate}
  \end{proposition}

  We compute the differential of $H$ implicitly using the relation \eqref{equ:dt_to_dx}. We will use the notation $\cross{v}:\real{3}\to\real{3\times 3}$ to denote the matrix representation of the cross product with the vector $v$, i.e.,
  \begin{equation}
    \bmat{v_1\\v_2\\v_3}\mapsto \bmat{0 & -v_3 &v_2\\v_3 & 0 & -v_1\\-v_2 & v_1 & 0},
  \end{equation}
  such that $[v]_\times w=v\times w$ for any $w\in\real{3}$. One can verify by direct computation the following property:
  \begin{equation}\label{eq:asymmetric to cross}
    wv\transpose-vw\transpose=\cross{\cross{v}w}.
  \end{equation}
  \begin{proposition}
    Let $\nu_\cF(t)$ represent a parametric curve. Then we have
    \begin{equation}
      \dot{H}=H\cross{-2M\dot{\nu}_{\cF}}
    \end{equation}
    where the matrix $M\in\real{3\times3}$ is given by
    \begin{equation}
      M=[u]_{\times}  \frac{ \left( I - u u\transpose \right) \left( I- \nu_\cF \nu_\cF\transpose \right)} {\norm{u'} \norm{\nu_\cF}}.
    \end{equation}
  \end{proposition}

\subsection{Ellipsoidal reachability constraints}\label{Chapter:ProblemFormulation}
  The reachability region is defined as the set of locations $x(t)$ that a robot can reach between two given fixed positions:

@@

\appendix

\section{Proof of proposition 1 }
  \begin{proof}
    For subclaim~\ref{it:orthonormality}:
    \begin{equation}
      H\transpose H=H^2=4uu\transpose u u\transpose - 4uu\transpose +I^2=I,
    \end{equation}
    since $u\transpose u=1$.

    For subclaim~\ref{it:determinant}, let $U=\bmat{u & u^\bot_{1},u^\bot_{2}}$, where $u^\bot_{1},u^\bot_{2}$ are two orthonormal vectors such that $I=UU\transpose=uu\transpose + u^\bot_{1}(u^\bot_{1})\transpose+u^\bot_{2}(u^\bot_{2})\transpose$; then, substituting $I$ in \eqref{eq:H definition}, we have that the eigenvalue decomposition of $H$ is given by
    \begin{equation}
      H=U\diag(1,-1,-1) U\transpose.
    \end{equation}
    Since the determinant of a matrix is equal to the product of the eigenvalues, $\det(H)=1$.

    For subclaim~\ref{it:transformation}, first note that $Hu=2uu\transpose u - u=u$.
    It follows that the sum of $\nu_\cF$ and $\nu_\cE$ is invariant under $H$:
    \begin{equation}\label{equ:H(u1+u2)}
      H(\nu_\cF+\nu_\cE)=H u \norm{\nu_\cF+\nu_\cE}
      = u\norm{\nu_\cF+\nu_\cE} =\nu_\cF+\nu_\cE,
    \end{equation}
    and that their difference is flipped under $H$:
    \begin{equation}\label{equ:H(u1-u2)}
      H(\nu_\cF-\nu_\cE) = 2uu\transpose (\nu_\cF-\nu_\cE) - (\nu_\cF-\nu_\cE)^2 = -(\nu_\cF-\nu_\cE).
    \end{equation}
    Combining (\ref{equ:H(u1+u2)}) and (\ref{equ:H(u1-u2)}) we obtain
    \begin{equation}
      H\nu_\cF = \frac{1}{2}\bigl(H(\nu_\cF+\nu_\cE)+H(\nu_\cF-\nu_\cE)\bigr)
      = \nu_\cE
    \end{equation}
  \end{proof}

\section{Proof or proposition 2}
\begin{proof}
    From the definition of $H$ in \eqref{eq:H definition}, we have
    \begin{equation} \label{equ:H_dot original}
      \dot H =   2(\dot u u\transpose + u \dot u\transpose)
    \end{equation}

    Recall that $\dot{u}=\frac{1}{\norm{u'}}(I-uu\transpose)\dot{u}'$ (see, for instance, \cite{Tron:Arxiv14}), which implies $(I-uu\transpose)\dot{u}'=\dot{u}'$. It follows that $\dot{u}$ flips sign under the action of $H\transpose$:
    \begin{multline}
      H\transpose \dot u = (2u u\transpose-I)\frac{ \left( I - u u\transpose \right) } {\left\|u'\right\|} \dot u' \\
      =\frac{1}{\norm{u'}} (2u u\transpose-I-2u u\transpose u u\transpose+u u\transpose)\dot{u}'\\
      = -\frac{1}{\norm{u'}} (I-u u\transpose)\dot u'
      = -\dot u
    \end{multline}

    Inserting $HH\transpose=I$ in (\ref{equ:H_dot original}), and using \ref{eq:asymmetric to cross}, we finally have
    \begin{equation}
      \begin{split}
        \dot H =  &  2H H\transpose(\dot u u\transpose + u \dot u\transpose)
        =  2 H (-\dot u u\transpose + u \dot u\transpose)\\
        =  &  -2H [[u]_{\times} \dot u]_{\times} \\
        = &  -2 H \left[ [u]_{\times}  \frac{ \left( I - u u\transpose \right) \left( I- \nu_\cF \nu_\cF\transpose \right)} {\left\|u'\right\| \left\|\nu_\cF\right\|} \dot{\nu}_\cF\right]_{\times}\\
        = & -2 H \cross{M \dot{\nu}_\cF},
      \end{split}
    \end{equation}
    which is equivalent to the claim.
  \end{proof}
{\small
\bibliographystyle{unsrt}
\bibliography{ADMM_planning,ACC}
}
